---
title: "TD_EM_Algorithm"
output: pdf_document
date: "2022-12-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Do Thanh Dat LE

Gan WANG

# TD à rendre sur l'agorithme EM

## Simulation

###1.

Simulate a sample of Poisson distribution with 100 observations and parameter $\lambda = 3$ 

```{r}
n1 <- 100
lambda1 <- 3
sample1 <- rpois(n1, lambda1)
sample1
```

###2.

Simulate a sample of Poisson distribution with 200 observations and parameter $\lambda = 15$ 

```{r}
n2 <- 200
lambda2 <- 15
sample2 <- rpois(n2, lambda2)
sample2
```

###3.

Create a vector with 300 integer value include 100 value "1" and followed by 200 value "2"

```{r}
z <- c(rep(1, 100), rep(2, 200))
table(z)
```

###4. Simulate a mixture of two Poisson distribution components

\[P(x)=\pi_1\dfrac{e^{-\lambda_1}}{x!}\lambda_1^{x}+\pi_2\dfrac{e^{-\lambda_2}}{x!}\lambda_2^{x}\] with $\lambda_1 = 3$ and $\lambda_2 = 15$, $\pi_1 = \dfrac{1}{3}$.

```{r}
N <- n1 + n2
z <- sample(z)
lambda <- c(3, 15)
mixture <- rpois(N, lambda = lambda[z])
mixture
hist(mixture, breaks=15)
```

# Algorithme EM pour une mélange de lois de Poisson à composantes

```{r}
# Create function for EM algorithm for mixture model of Poisson distribution
EM.algorithm <- function(X, K, lambda_0, epsilon) {
  # EM Initialisation
  pi_0 <- rep(1/K,K)
  pi_q <- pi_0
  lambda_q <- lambda_0
  T <- matrix(NA, nrow = length(X), ncol = K)
  while (TRUE) {
    theta_q <- c(pi_q, lambda_q)
    # E step
    for (k in 1:K) {
      T[,k] <- dpois(X, lambda = lambda_q[k])
    }
    for (k in 1:K) {
      T[,k] <- pi_q[k]*T[,k]
    }
    T <- T / rowSums(T)
    # M step
    pi_new <- colSums(T)/sum(T)
    lambda_new <- (X%*%T)/colSums(T)
    theta_new <- c(pi_new, lambda_new)
    delta = (norm(as.matrix(theta_new)-as.matrix(theta_q), 
            type = "2"))^2/(norm(as.matrix(theta_q), type = "2"))^2
    if (delta < epsilon) {
      break
    } else {
      pi_q <- pi_new
      lambda_q <- lambda_new
    }
  }
  return(theta_new)
}
```

#Test the EM algorithm
```{r}
result1 <- EM.algorithm(X = mixture, K = 2, lambda_0 = c(5,10), epsilon = 10^(-8))
pi1 <- result1[1]
pi2 <- result1[2]
lambda1 <- result1[3]
lambda2 <- result1[4]
pi1
pi2
lambda1
lambda2
```